SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-20 paper/intro_MTedit.tex
more in intro

The analysis of VTSs is a difficult computational problem because of the combinatorial scaling of graph topologies and regulatory rules. Many questions require us to check useful properties over all possible graphs and regulatory rules. Formal verification tools such as model checkers \cite{clarke1996symbolic, biere2003bounded, clarke2008birth, cimatti2000nusmv, holzmann1997model} and SAT solvers \cite{moskewicz2001chaff,een2004extensible} allow us to do this symbolically for graphs of finite size, without enumerating all instances. In a recent work, we employed the model checker CBMC \cite{CKY03, ckl2004} meant to analyze ANSI C programs for studying this problem. VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. Queries about graph connectivity requirements for different variations of SNARE regulations were modeled as logical assertions to be checked by the model checker. CBMC uses its own built-in encodings to reduce the analysis problem to a Boolean satisfiability problem which is solved using SAT solvers.
%
However, the scalability of our recent approach was limited (up to
VTSs with 6 nodes) due to the encoding used to model VTSs was not
fine-tuned to the structure of the problem.
%
In particular, we needed to encode conditional reachability between
nodes (refer section 3, stability condition), and solved using a
combination of non-determinism and partial enumeration that is not as
efficient.

@@

% \todo{REQUIRE A REWRITE FLOW PROBS}
In this paper, we have developed a novel SAT encoding for
searching VTSs that satisfy the given properties.
%
We have improved the encoding of some of the conditions for VTSs.
%
Especially, we have recursively defined the reachability condition
such that that we avoid the inefficient enumeration.
%
% We have also encoded $k$-connectivity in a similar fashion and avoided
% a blowup in the formula size.
%
Furthermore, our encoding supports several variations of the properties.
%
We have implemented the encoding using Z3~\cite{z3} python API and
searched for VTS satisfying various properties upto size 14-18 nodes
as compare earlier tool that could only search graph upto size 6
nodes.
%
Our tool is useful to cell biologists since VTSs of real cells have
approximately ten compartments.

>>>>

The analysis of VTSs is a difficult computational problem because of the combinatorial scaling of graph topologies and regulatory rules. Many questions require us to check useful properties over all possible graphs and regulatory rules. Formal verification tools such as model checkers \cite{clarke1996symbolic, biere2003bounded, clarke2008birth, cimatti2000nusmv, holzmann1997model} and SAT solvers \cite{moskewicz2001chaff,een2004extensible} allow us to do this symbolically for graphs of finite size, without enumerating all instances. In a recent work, we employed the model checker CBMC \cite{CKY03, ckl2004} meant to analyze ANSI C programs for studying this problem. VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. Queries about graph connectivity requirements for different variations of SNARE regulations were modeled as logical assertions to be checked by the model checker. CBMC uses its own built-in encodings to reduce the analysis problem to a Boolean satisfiability problem which is solved using SAT solvers.
%
However, the scalability of our recent approach was limited (up to
VTSs with 6 nodes) due to the encoding used to model VTSs was not
fine-tuned to the structure of the problem.
%
In particular, we needed to encode conditional reachability between
nodes (refer section 3, stability condition), and solved in CBMC using a
combination of non-determinism and partial enumeration that is
inefficient.

@@

% \todo{REQUIRE A REWRITE FLOW PROBS}
In this paper, we have developed a novel SAT encoding for
searching VTSs that satisfy the useful properties.
%
We have improved the encoding of some of the conditions for VTSs.
%
Especially, we have recursively defined the reachability condition
such that that we avoid the inefficient enumeration.
%
% We have also encoded $k$-connectivity in a similar fashion and avoided
% a blowup in the formula size.
%
% Furthermore, our encoding supports several variations of the properties.
%
We have implemented the encoding using Z3~\cite{z3} python API and
searched for VTS satisfying various properties upto size 14-18 nodes
as compare earlier tool that could only search graph upto
size~6\footnote{The experiments in~\cite{shukla} scale upto 8.
However, our experiments only scale upto 6, because
we are using less powerful machine.} nodes.
%
Our tool is useful to cell biologists since VTSs of real cells have
approximately 10 compartments.
